---
title: "toolik_chem"
author: "OP"
format: html
editor: visual
execute: 
  echo: false
  warning: false
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(purrr)
```

```{r}
toolik_biochem <- read_csv(here("data","2011_Kling_Akchem.csv"),na=".") %>% 
  clean_names()
```

Create a subset of the data that contains only observations from the "Toolik Inlet" site, and that only contains the variables (columns) for pH, dissolved organic carbon (DOC), and total dissolved nitrogen (TDN) (**hint:** see `dplyr::select()`). Store this subset as `inlet_biochem`. Make sure to look at the subset you've created.

```{r}

inlet_biochem <- toolik_biochem %>% 
  dplyr::select(p_h,doc_u_m )
```

Find the mean value of each column in inlet_biochem 3 different ways:

Write a for loop from scratch to calculate the mean for each Use one other method (e.g. apply, across, or purrr::map_df) to find the mean for each column.

```{r}

mean_inlet_biochem <- apply(inlet_biochem, 2, mean, na.rm = TRUE)
mean_inlet_biochem
```

Write a for loop from scratch to calculate the mean for each

```{r}

for(i in seq_along(inlet_biochem)){
  mean_inlet_biochem_for_loop <- mean(inlet_biochem[[i]], na.rm = TRUE)
  print(mean_inlet_biochem_for_loop)
}
```

Use *one other method* (e.g. `apply`, `across`, or `purrr::map_df`) to find the mean for each column.

```{r}
# use accross function to find mean
 inlet_biochem %>% 
    summarize(across(everything(), mean, na.rm = TRUE))

```

```{r}

# use purrr::map_df function to find mean

 inlet_biochem %>% 
    map_df(mean, na.rm = TRUE)
  
```
